Paper

### 2020-01-10 Paper_resubmit.Rmd
...

@@

@@

@@

```{r, eval=F}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
load(file, verbose=T)

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
v.tiles = do.call(bind_rows, purrr::map(list.files(vdir, '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.tiles = filter(v.tiles, Samplename %in% val.info$Samplename)
v.tiles = v.tiles[-which(duplicated(v.tiles$Samplename))[2],]

v.arm.tiles = do.call(bind_rows, purrr::map(list.files(vdir, 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.arm.tiles = filter(v.arm.tiles, Samplename %in% val.info$Samplename)
v.arm.tiles = v.arm.tiles[-which(duplicated(v.arm.tiles$Samplename))[2],]

for (i in 1:length(z.mean))
  v.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.tiles[,i+1], z.mean[i], z.sd[i])

for (i in 1:length(z.arms.mean))
  v.arm.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.arm.tiles[,i+1], z.arms.mean[i], z.arms.sd[i])

vt = dplyr::select(v.tiles, -Samplename) %>% as.matrix; rownames(vt) = v.tiles$Samplename
at = dplyr::select(v.arm.tiles, -Samplename) %>% as.matrix; rownames(at) = v.arm.tiles$Samplename

v.df = BarrettsProgressionRisk:::subtractArms(vt, at)

valCNdf = as_tibble(t(v.df), rownames='locs') %>% set_names(c('locs',val.info$Samplename)) %>%
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>%   
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  #dplyr::mutate( seg.type = '5MB' ) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(val.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')'))

#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value)) + ylim(-30,20)

@@

residuals = do.call(bind_rows, purrr::map(list.files( dirname(vdir), 'residuals.tsv', full.names = T, recursive = T ), function(f) {
  read_tsv(f, col_types = 'ccddddl')
}))

failedQC = residuals %>% filter(varMAD_median > 0.008)

@@

```{r, fig.width=12, fig.height=8}
model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
d.n.patient = length(unique(pg.samp$Patient))
d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds', transpose=T)

@@

load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
dv.coefs = coefs[['0.9']]
dv.perf = performance.at.1se[['0.9']]$mean
rm(dysplasia.df, plots, pg.samp, models, nzcoefs, models, performance.at.1se)

cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
filenames = list.files(cache.dir, 'supp*')

info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
  load(f,verbose=F)
  model.info
})) %>% mutate('Cohort' = 'Discovery (patient subsets)')

info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)')

info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation')

x = info %>% group_by(n.patients, Cohort) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation'), ordered=T))

p = ggplot(x, aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
  scale_fill_brewer(palette = 'Set2',name='') + 
  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='Cross validation performance', y='mean', x='# patients used to train')

@@

p1 = ggplot(x, aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
  scale_y_continuous(expand=c(0,0), limits=c(0,125)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='# Coefficients',y='mean', x='# patients used to train')

p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='AUC', y='mean', x='# patients used to train')

>>>>

error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')

@@

error = error %>% left_join(pg.samp %>% dplyr::select(Status, Samplename, RR), by=c('Sample'='Samplename')) %>% 
  dplyr::mutate(P.High = BarrettsProgressionRisk:::pi.hat(RR+Error), P.Low = BarrettsProgressionRisk:::pi.hat(RR-Error), Abs.Error = P.High-P.Low )

error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
summary(error$Abs.Error)
sd(error$Abs.Error)

@@

pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

@@

@@

@@

```{r, echo=F}
# SNP
snp.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
  mutate( PID = paste(PatientID, `Timepoint Code`, sep='_'), Patient = as.numeric(PatientID), Hospital.Research.ID = as.character(Patient), Samplename = PID )

model.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/'
load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))

dvsnp.coefs = coefs[['0.9']]
dvsnp.perf = performance.at.1se[['0.9']]$mean
rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)

dvsnp.preds = purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/loo', full.names = T), function(f) read_tsv(f, col_types = 'ccccccdd') )

dvsnp.preds = do.call(bind_rows, dvsnp.preds)
dvsnp.patient = length(unique(dvsnp.preds$Patient))
dvsnp.roc = pROC::roc(Status ~ Probability, data=dvsnp.preds, ci=T, of='thresholds', transpose=T)

dvsnp.dv.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort != 'snp') )
dvsnp.snp.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort == 'snp') )

dvsnp.preds = dvsnp.preds %>% mutate(type = ifelse(cohort == 'snp', 'snp', 'sWGS'))

dvsnp.aucs = left_join(dvsnp.preds %>% dplyr::select(Patient, type) %>% distinct %>% group_by(type) %>% tally(name='# patients'),
            dvsnp.preds %>% dplyr::select(Patient, type) %>% group_by(type) %>% tally(name='# timepoint samples'), by='type') %>% 
  add_row( type = 'All (snp+sWGS)', `# patients` = dvsnp.preds %>% dplyr::select(Patient) %>% distinct %>% tally %>% pull, `# timepoint samples` = dvsnp.preds %>% dplyr::select(Patient) %>% tally %>% pull) %>% 
  arrange(type) %>% add_column( auc = c(dvsnp.roc$auc, dvsnp.snp.auc, dvsnp.dv.auc) ) %>%
  mutate_at(vars(-auc), as.factor)

dv.snp.p = ggplot(dvsnp.aucs , aes(`# patients`, auc, group=type, fill=type)) + geom_col() + 
  geom_text(aes(label=round(auc,2), y=0.05)) + 
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2', name='') + 
  labs(title='sWGS (all) + SNP Model', y='AUC') + theme(legend.position = 'bottom') +
  plot.theme
dv.snp.p

ggsave(filename='plots/performance/training_snpdv.png', plot=dv.snp.p, width=6, height=10, units='in', dpi=300)

# discovery + validation
model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
d.n.patient = length(unique(pg.samp$Patient))
d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds', transpose=T)

@@

load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
dv.coefs = coefs[['0.9']]
dv.perf = performance.at.1se[['0.9']]$mean
rm(dysplasia.df, plots, pg.samp, models, nzcoefs, cvs, performance.at.1se)
```

```{r, fig.width=12, fig.height=8}
cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
filenames = list.files(cache.dir, 'supp*')

info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
  load(f,verbose=F)
  model.info
})) %>% mutate('Cohort' = 'Discovery (patient subsets)', Type = 'sWGS')

info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)', Type = 'sWGS')

info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation', Type = 'sWGS')

info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc, Cohort = 'sWGS (all) + SNP', Type = 'All')

info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'SNP')

info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'sWGS')

x = info %>% group_by(n.patients, Cohort, Type) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients,levels=c( seq(28,88,10), 164, 413, 165, 248  ), ordered=T), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation', 'sWGS (all) + SNP'), ordered=T))

p = ggplot( x %>% filter(!is.na(cv.performance_mean)), aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
  scale_fill_brewer(palette = 'Set2',name='') + 
  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='Cross validation performance', y='', x='# patients used to train')

@@

p1 = ggplot(x %>% filter(!is.na(n.coefs_mean)), aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
  scale_y_continuous(expand=c(0,0)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='# Coefficients',y='', x='# patients used to train')

p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
  geom_col() + 
  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
  geom_text( data=filter(x,Cohort == 'sWGS (all) + SNP'), aes(label=Type, y=0.1) ) +
  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
  labs(title='AUC', y='', x='# patients used to train')
